### 2018-01-30 CO749-GraphColoring/assignments/A2.tex
first two claims of 3

\section{}
\paragraph{}
We opt to prove the BONUS Lemma, i.e. Lemma $3$:
\textit{Let $G$ be a graph and for each vertex $v$ of $G$ let $r(v)$ be a non-negative integer. If $L$ is a list assignment of $G$ such that for every vertex $v$ of $G$
$$|L(v)| \geq \frac{d(v) + 1}{r(v)+1},$$
then $G$ has an $r(v)$-defective coloring}. This Lemma directly implies Lemma $2$ of this problem.
\paragraph{}
Consider a graph $G$, non-negative integers $r(v)$ for each $v \in V(G)$, and a list assignment $L$ of $G$ such that for each $v \in V(G)$
$$|L(v)| \geq \frac{d(v) + 1}{r(v)+1}.$$
Let $\phi$ be an improper $L$-coloring of $G$ minimizing the number of ``improper" edges $uv \in E(G)$ such that $\phi(u) = \phi(v)$. Suppose for a contradiction that $\phi$ is not an $r(v)$-defective coloring.
\paragraph{}
Then there exists a vertex $v$ such that the number of neighbours of $v$ receiving the same color as $v$ under $\phi$ is at least $r(v) + 1$. If there exists a color $c \in L(v)$ such that the number of neighbours of $v$ colored $c$ is at most $r(v)$ then we can construct a new coloring $\phi'$ by setting $\phi'(v) = c$ and $\phi'(u) = \phi(u)$ for all $u \neq v$. This coloring has $r(v)$ new improper edges, those between $v$ and its $c$-colored neighbours. But $\phi'$ switches $r(v) + 1$ edges which were improper under $\phi$, those between $v$ and its $\phi(v)$-colored neighbours, to proper edges. Hence $\phi'$ has fewer improper edges than $\phi$, a contradiction.
\paragraph{}
If such a color $c \in L(v)$ does not exist then  for each $c \in L(v)$ the number of neighbours of $v$ colored $c$ under $\phi$ is at least $r(v) + 1$. But then $v$ has at least
$$|L(v)|(r(v) + 1) \geq d(v) + 1$$
neighbours, a contradiction. Therefore $\phi$ is an $r(v)$-defective coloring $\blacksquare$
\end{document}

>>>>

\newpage
\section{}
\paragraph{}
We opt to prove the BONUS Lemma, i.e. Lemma $3$:
\textit{Let $G$ be a graph and for each vertex $v$ of $G$ let $r(v)$ be a non-negative integer. If $L$ is a list assignment of $G$ such that for every vertex $v$ of $G$
$$|L(v)| \geq \frac{d(v) + 1}{r(v)+1},$$
then $G$ has an $r(v)$-defective coloring}. This Lemma directly implies Lemma $2$ of this problem.
\paragraph{}
Consider a graph $G$, non-negative integers $r(v)$ for each $v \in V(G)$, and a list assignment $L$ of $G$ such that for each $v \in V(G)$
$$|L(v)| \geq \frac{d(v) + 1}{r(v)+1}.$$
Let $\phi$ be an improper $L$-coloring of $G$ minimizing the number of ``improper" edges $uv \in E(G)$ such that $\phi(u) = \phi(v)$. Suppose for a contradiction that $\phi$ is not an $r(v)$-defective coloring.
\paragraph{}
Then there exists a vertex $v$ such that the number of neighbours of $v$ receiving the same color as $v$ under $\phi$ is at least $r(v) + 1$. If there exists a color $c \in L(v)$ such that the number of neighbours of $v$ colored $c$ is at most $r(v)$ then we can construct a new coloring $\phi'$ by setting $\phi'(v) = c$ and $\phi'(u) = \phi(u)$ for all $u \neq v$. This coloring has $r(v)$ new improper edges, those between $v$ and its $c$-colored neighbours. But $\phi'$ switches $r(v) + 1$ edges which were improper under $\phi$, those between $v$ and its $\phi(v)$-colored neighbours, to proper edges. Hence $\phi'$ has fewer improper edges than $\phi$, a contradiction.
\paragraph{}
If such a color $c \in L(v)$ does not exist then  for each $c \in L(v)$ the number of neighbours of $v$ colored $c$ under $\phi$ is at least $r(v) + 1$. But then $v$ has at least
$$|L(v)|(r(v) + 1) \geq d(v) + 1$$
neighbours, a contradiction. Therefore $\phi$ is an $r(v)$-defective coloring $\blacksquare$

\newpage
\section{}
\paragraph{}
We prove a slightly more general result allowing $m(v)$ to take the value $0$. This will make induction easier when we wish to delete from a list of size $2$. For such a $v$, $L(v)= \emptyset$ and the desired $S(v) = \emptyset$. Everything else is exactly as stated in the problem.
\paragraph{}
Suppose not. Let $G$ be a counterexample satisfying:
\begin{enumerate}
	\item $\sum_{v \in V(G)} m(v)$ is minimized, and subject to that
	\item if possible $G(c)$ is connected for each $c \in L(G)$, and subject to that possibility
	\item $|L(G)|$ is minimized.
\end{enumerate}
\begin{claim}
	\begin{enumerate}
		\item For all $c \in L(G)$, $G(c)$ is connected.
		\item For every pair of distinct colors $c,c' \in L(G)$, $G(c) \neq G(c')$.
		\item For every pair of distinct colors $c, c' \in L(G)$, if $V(G(c)) \cap V(G(c')) = \emptyset$ then $G[V(G(c))\cup V(G(c'))]$ is disconnected.
	\end{enumerate}
\end{claim}
\begin{proof}
	\paragraph{} For (1.). While there exists a color $c \in L(G)$ such that $G(c)$ has at least two connected components $G_1$ and $G_2$ do the following:
		\subparagraph{}
		Let $c'$ be a new color not contained in $L(G)$. For each $v \in V(G_2)$ let $$L'(v) = (L(v)\cup \{c'\}) \backslash \{c\}.$$ For each $v \in V(G) \backslash V(G_2)$ let $L'(v) = L(v)$. Replace $L$ with $L'$.
	\paragraph{}
	We claim that every iteration of the above preserves the property of $(G,m,L)$ being a counterexample, but decreases the ``disconnectivity". i.e. the sum over $c \in L(G)$ such that $G(c)$ is disconnected, of the number of connected components in $G(c)$.
	\paragraph{}
	It is clear the disconnectivity decreases. We add a new color $c'$ for which $G(c')$ is connected, and $G(c)$ loses one connected component, and the number of connected components of $G(c'')$ is not affected for any other color $c'' \in L(G)$.
	\paragraph{}
	To verify that the new lists, say $L'$, also admit a counterexample with $G$ and $m$, suppose to contrary there are $S'(v) \subseteq L'(v)$ of size $m(v)$ for each $v \in V(G)$ and $S'(u) \cap S'(v) = \emptyset$ for all $uv \in E(G)$. For each $v \in V(G)$ let $$S(v) = \begin{cases} S'(v)\cup \{c\} \backslash \{c'\}, &\text{if } v \in V(G_2) \text{ and } c' \in S'(v) \\
	S'(v), &\text{otherwise}.\end{cases} $$
	Then $S(v) \subseteq L(v)$ and $|S(v)| = m(v)$ for each $v \in V(G)$. Further, for each $uv \in E(G)$, $S(u) \cap S(v) = \emptyset$. To see this, observe that if $S(u) \cap S(v) \neq \emptyset$ then by our choice of $S'$, $S(u) \cap S(v) = \{c\}$ and, since $G(c)$ was disconnected, $u, v \in V(G_2)$. But then, $S'(u) \cap S'(v) = \{c'\}$, a contradiction.
	\paragraph{}
	So the above algorithm will terminate when the ``disconnectivity" reaches $0$ and we will have found a connected counterexample with the same value for $\sum_{v} m(v)$ as $G$. This contradicts our choice of $G$ if there exists color $c$ such that $G(c)$ is disconnected. Therefore $G(c)$ is connected for all $c \in L(G)$.
	\paragraph{}For (2.). Suppose for a contradiction there exist distinct colors $c,c' \in L(G)$ such that $G(c) = G(c')$. Construct new lists $L'$ such that for each $v \in V(G)$
	$$ L'(v) = \begin{cases}
	L(v) \backslash \{c,c'\}, &\text{ if } v \in V(G(c)) \\
	L(v), &\text{otherwise.}
		\end{cases}$$
	Let $m'(v) = m(v) - 1$ if $v \in V(G(c))$ (this is well-defined since such $m(v) \geq 2$) and let $m'(v) = m(v)$ otherwise.  By the minimality of $G$, there exist for each $v \in V(G)$, a set $S'(v)\subseteq L'(v)$ of size $m'(v)$. These sets satisfy $S'(u) \cap S'(v) = \emptyset$ for all $uv \in E(G)$. Order the vertices of $G(c)$ as $v_1, \dots, v_k$ with respect to some orientation of $G$. Construct sets $S(v)$ as follows. If $v \in V(G) \backslash V(G(c))$ set $S(v) = S'(v)$. If $v_i \in V(G(c))$ and $i$ is even set $S(v) = S'(v) \cup \{c\}$. If $v_i \in V(G(c))$ and $i$ is odd set $S(v) = S'(v) \cup \{c'\}$. Since $c$ and $c'$ do not appear in sets outside $G(c)$, and we alternate $c$ and $c'$ along $G(c)$, we have $S(u) \cap S(v) = \emptyset$ for all $uv \in E(G)$. But this contradicts that $(G,m,L)$ is a counterexample. Hence for distinct colors $c$ and $c'$, $G(c) \neq G(c')$.

\end{proof}

\end{document}
